---
title: "ihs5_cleaning2"
author: "Gareth Osman"
date: "5/11/2021"
output: html_document
---

INTRO 

1) We need to identify where the data on household consumption is stored. 

For ihs5, the data on food consumption was recorded as part of the household questionnaire (datasets starting w/ HH_). Within this questionnaire, data on food consumption is recorded in the module G (Food consumption). Within the module G, we find 3 sets of data, each one with different level of aggregation. We are interested in g1 which is food recorded at food item level (disaggregated data). 

Hence, we need to load the file hh_mod_g1 (household questionaire module G, 1- consumption at item level). Once it is loaded, we can check that the file is loaded correctly by checking that the observations (cases) and variables match the one reported in the worldbank website (https://microdata.worldbank.org/index.php/catalog/3818)

```{r}
#--- Load libraries 
library(plyr)
library(haven)
library(magrittr)
library(foreign)
library(psych)
library(readxl)
library(labelled)
library(tidyverse)

#--- Change to show no scientific notation & round to 3 decimal places
options(scipen = 10, digits=3) 

#--- Read in the file
ihs5 <- haven::read_dta(here::here('HH_MOD_G1.dta'))

```

>CHANGE VARIABLE NAMES AND DATA CLASS 

The next step, the variables must then be standardized. This is important because we will be using data from various countries, and the variable names will vary from questionnaire to questionnaire. However, if we want to standardize (or harmonize) our cleaning and processing structure, it is best to standardize variable names (when possible). Within the data dictionary we can find a description of each variable. And, below is the codebook with the new variable names.

Variable category description: (i) hh_g03 refer to total consumption 
                               (ii) hh_g04 refers to purchased by the household 
                               (iii) hh_06 refers to food from own-production 
                               (iv) hh_g07 refers to food consumed from gift (and                                         other sources). 

Then, to identify variables that report quantity will be those that has an 'a' after the number (i.e. hh_g03a), variables reporting units will be identified by 'b' (i.e. hh_g07b, or hhg07b_label, hg07b_oth) and 'c' is for sub-unit related to photo aid guidance. 

Codebook (https://microdata.worldbank.org/index.php/catalog/3818/data-dictionary)

**Variable names need to be corrected!!** gift and purchased have the same code


```{r}

#--- Renaming variables to standard names 

#Summary info 
 ihs5 <- ihs5 %>% rename(
      id = 'case_id',
      consYN = 'hh_g01',
      item_code = 'hh_g02',
      item_oth = 'hh_g01_oth')


ihs5$id <- as.character(ihs5$id)
ihs5$consYN <- as.factor(ihs5$consYN)
ihs5$item_name <- as.factor(ihs5$item_code)
ihs5$item_oth <- as.factor(ihs5$item_oth)


#G03: Quantity and units consumed of each food 
ihs5 <- ihs5 %>% rename(
      cons_quant = 'hh_g03a',
      cons_unitA = 'hh_g03b',
      cons_unit = 'hh_g03b_label',
      cons_unit_oth = 'hh_g03b_oth',
      cons_unit_size = 'hh_g03c')

ihs5$cons_quant <- as.numeric(ihs5$cons_quant)
ihs5$cons_unitA <- as.character(ihs5$cons_unitA)
ihs5$cons_unit <- as.character(ihs5$cons_unit)
ihs5$cons_unit_oth <- as.factor(ihs5$cons_unit_oth)
ihs5$cons_unit_size <- as.factor(ihs5$cons_unit_size)


#G04: Quantity and units purchased of each food 
ihs5 <- ihs5 %>% rename(
      purc_quant = 'hh_g04a',
      purc_unitA = 'hh_g04b',
      purc_unit = 'hh_g04b_label',
      purc_unit_oth = 'hh_g04b_oth',
      purc_unit_size = 'hh_g04c',
      purc_price = 'hh_g05')

ihs5$purc_quant <- as.numeric(ihs5$purc_quant)
ihs5$purc_unitA <- as.factor(ihs5$purc_unitA)
ihs5$purc_unit <- as.factor(ihs5$purc_unit)
ihs5$purc_unit_oth <- as.factor(ihs5$purc_unit_oth)
ihs5$purc_unit_size <- as.factor(ihs5$purc_unit_size)
ihs5$purc_price <- as.numeric(ihs5$purc_price)

#G06: Quantity and units home produced of each food
ihs5 <- ihs5 %>% rename(
      prod_quant = 'hh_g06a',
      prod_unitA = 'hh_g06b',
      prod_unit = 'hh_g06b_label',
      prod_unit_oth = 'hh_g06b_oth',
      prod_unit_size = 'hh_g06c')

ihs5$prod_quant <- as.numeric(ihs5$prod_quant)
ihs5$prod_unitA <- as.factor(ihs5$prod_unitA)
ihs5$prod_unit <- as.factor(ihs5$prod_unit)
ihs5$prod_unit_oth <-as.factor(ihs5$prod_unit_oth)
ihs5$prod_unit_size <- as.factor(ihs5$prod_unit_size)

#G07: Quantity and units gifted/other source of each food 
ihs5 <- ihs5 %>% rename(
      gift_quant = 'hh_g07a',
      gift_unitA = 'hh_g07b',
      gift_unit = 'hh_g07b_label',
      gift_unit_oth = 'hh_g07b_oth',
      gift_unit_size = 'hh_g07c')

ihs5$gift_quant <- as.numeric(ihs5$gift_quant)
ihs5$gift_unitA <- as.factor(ihs5$gift_unitA)
ihs5$gift_unit <- as.factor(ihs5$gift_unit)
ihs5$gift_unit_oth <- as.factor(ihs5$gift_unit_oth)
ihs5$gift_unit_size <- as.factor(ihs5$gift_unit_size)

#Removing a duplicate variable- item_code
 ihs5 <- select(ihs5, -33)
 
```

>EXTRACTING VALUES AND LABELS OF LABELED 'ITEM_CODE'

The ihs5 dataset has stored the variable 'item_code' as an attribute (dbl+lbl). Hence we need to extract a two column data frame of value and label. The label and values are stored as an attribute of data frame. 

```{r}
ihs5 %>% distinct(item_code)

ihs5_attr <- stack(attr(ihs5$item_code, 'labels')) 

#Changing variable class
ihs5_attr$values <- as.factor(ihs5_attr$values)
ihs5$item_code <- as.factor(ihs5$item_code)

#Merging the values and labels back into the ihs5 dataset
ihs5 %>% 
    left_join(., ihs5_attr, by=c("item_code" = "values"))
ihs5 <- ihs5 %>% 
     left_join(., ihs5_attr, by=c("item_code" = "values"))

ihs5 <- ihs5 %>% relocate("ind", .after = item_code)
ihs5 <- ihs5 %>% rename(item_name_original = 'ind')

ihs5 <- ihs5 %>% arrange(cons_unit) %>% arrange(item_code) %>% arrange(HHID)
ihs5 <- ihs5 %>% select(id:consYN, item_code, item_name_original, item_oth:cons_unitA, cons_unit, cons_unit_oth:gift_unit_size)

```

>CLEANING UNITS LABELED AS 'OTHER'

Food items and units are coming from a standard list. However, some food consumed by the household is not in that list of items, hence it can be recorded under 'other'. The code to identify 'other food items varies depending on the food group

Similarly, for units, when the household is reporting the food consumed in a unit that it is not in the list will be recoded and the unit code apply for 'other' is 23.

There are a number of item units recorded as "23" but described as existing ucodes in our conversion factor file. We need to change these units from "23" to their appropriate units as it is in the 2020 cleaned conversion factor file. 
Any item without a standard unit will be excluded in the analysis. 

```{r}
ihs5 %>% filter(cons_unitA==23)

ihs5 %>% filter(cons_unit_oth!= "") %>% count(cons_unit_oth) %>% arrange(desc(n)) %>% print(n = nrow(ihs5))

#units listed as "other" and have not been converted to standard units
ihs5 %>% filter(cons_unitA==23) %>% count(cons_unit_oth) %>% arrange(desc(n)) %>% print(n = nrow(ihs5))

ihs5 %>% filter(., cons_unit_oth=="HEAP") %>% select(item_name_original, item_code, cons_quant, cons_unitA, cons_unit, cons_unit_oth) %>% arrange(item_name_original) %>% print(n = nrow(ihs5))

ihs5 %>% filter(item_code=="803") %>% select(item_name_original, item_code, cons_quant, cons_unitA, cons_unit, cons_unit_oth) %>% arrange(desc(cons_unitA))  %>% print(n = nrow(ihs5))

```


```{r}
ihs5[ihs5$cons_unit_oth == "PACKET", "cons_unitA"] <- '51'
ihs5[ihs5$cons_unit_oth == "PACKET UNSPECIFIED", "cons_unitA"] <- '51'
ihs5[ihs5$cons_unit_oth == "TABLESALT PACKET", "cons_unitA"] <- '51'
ihs5[ihs5$cons_unit_oth == "PACKET MEDIUM", "cons_unitA"] <- '51'
ihs5[ihs5$cons_unit_oth == "MEDIUM PACKET", "cons_unitA"] <- '51'
ihs5[ihs5$cons_unit_oth == "SATCHET", "cons_unitA"] <- '22'
ihs5[ihs5$cons_unit_oth == "TUB", "cons_unitA"] <- '22'
ihs5[ihs5$cons_unit_oth == "TBE", "cons_unitA"] <- '22'
ihs5[ihs5$cons_unit_oth == "SA HETS", "cons_unitA"] <- '22'
ihs5[ihs5$cons_unit_oth == "TINA", "cons_unitA"] <- '25'
ihs5[ihs5$cons_unit_oth == "GRAMS", "cons_unitA"] <- '18'
ihs5[ihs5$cons_unit_oth == "G", "cons_unitA"] <- '18'
ihs5[ihs5$cons_unit_oth == "PACKET (SMALL)", "cons_unitA"] <- '54'
ihs5[ihs5$cons_unit_oth == "SMALL PACKET", "cons_unitA"] <- '54'
ihs5[ihs5$cons_unit_oth == "SMALL PACKAGE", "cons_unitA"] <- '54'
ihs5[ihs5$cons_unit_oth == "SMALL POCKET", "cons_unitA"] <- '54'
ihs5[ihs5$cons_unit_oth == "SMALL  PACKET", "cons_unitA"] <- '54'
ihs5[ihs5$cons_unit_oth == "PIECE", "cons_unitA"] <- '9'
ihs5[ihs5$cons_unit_oth == "MANGO", "cons_unitA"] <- '9'
ihs5[ihs5$cons_unit_oth == "BUNCH", "cons_unitA"] <- '8'
ihs5[ihs5$cons_unit_oth == "BATCHES", "cons_unitA"] <- '8'
ihs5[ihs5$cons_unit_oth == "BUNDLE", "cons_unitA"] <- '8'
ihs5[ihs5$cons_unit_oth == "MILLILITRE", "cons_unitA"] <- '19'
ihs5[ihs5$cons_unit_oth == "5L BUCKET", "cons_unitA"] <- '26'
ihs5[ihs5$cons_unit_oth == "PAIL (SMALL)", "cons_unitA"] <- '4A'
ihs5[ihs5$cons_unit_oth == "TABLE SPOON", "cons_unitA"] <- '59'
ihs5[ihs5$cons_unit_oth == "PAIL", "cons_unitA"] <- '4'
ihs5[ihs5$cons_unit_oth == "KILOGRAM", "cons_unitA"] <- '1'
ihs5[ihs5$cons_unit_oth == "PAIL (LARGE)", "cons_unitA"] <- '4C'
ihs5[ihs5$cons_unit_oth == "PAIL (MEDIUM)", "cons_unitA"] <- '4B'
ihs5[ihs5$cons_unit_oth == "LITRE", "cons_unitA"] <- '15'
ihs5[ihs5$cons_unit_oth == "TINA FLAT", "cons_unitA"] <- '25A'
ihs5[ihs5$cons_unit_oth == "SMALL TINA FLAT", "cons_unitA"] <- '25A'
ihs5[ihs5$cons_unit_oth == "SATCHET (SMALL)", "cons_unitA"] <- '22A'
ihs5[ihs5$cons_unit_oth == "SMALL SACHET", "cons_unitA"] <- '22A'
ihs5[ihs5$cons_unit_oth == "SMALL SARCHET", "cons_unitA"] <- '22A'
ihs5[ihs5$cons_unit_oth == "TINA HEAPED", "cons_unitA"] <- '25B'
ihs5[ihs5$cons_unit_oth == "HEAP TINA", "cons_unitA"] <- '25B'
ihs5[ihs5$cons_unit_oth == "TINA (HEAPED)", "cons_unitA"] <- '25B'
ihs5[ihs5$cons_unit_oth == "CLUSTER", "cons_unitA"] <- '44'
ihs5[ihs5$cons_unit_oth == "PACKET LARGE", "cons_unitA"] <- '55'
ihs5[ihs5$cons_unit_oth == "BIG PACKET", "cons_unitA"] <- '55'
ihs5[ihs5$cons_unit_oth == "SMALL HEAP", "cons_unitA"] <- '10A'
ihs5[ihs5$cons_unit_oth == "HEAP(SMALL)", "cons_unitA"] <- '10A'
ihs5[ihs5$cons_unit_oth == "LARGE HEAP", "cons_unitA"] <- '10C'
ihs5[ihs5$cons_unit_oth == "SMALL TIN", "cons_unitA"] <- '71'
ihs5[ihs5$cons_unit_oth == "MEDIUM SACHET", "cons_unitA"] <- '22B'
ihs5[ihs5$cons_unit_oth == "MEDIAM SACHET", "cons_unitA"] <- '22B'
ihs5[ihs5$cons_unit_oth == "BUNCH SMALL", "cons_unitA"] <- '8A'
ihs5[ihs5$cons_unit_oth == "LARGE  SATCHET", "cons_unitA"] <- '22C'
ihs5[ihs5$cons_unit_oth == "SACHET  LARGE", "cons_unitA"] <- '22C'
ihs5[ihs5$cons_unit_oth == "HEAP (MEDIUM)", "cons_unitA"] <- '10B'
ihs5[ihs5$cons_unit_oth == "HEAP", "cons_unitA"] <- '10'
ihs5[ihs5$cons_unit_oth == "50 GRAM", "cons_unitA"] <- '42'

ihs5[ihs5$item_code==404 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==829 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==101 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==205 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==821 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==107 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==302 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6B'
ihs5[ihs5$item_code==305 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6B'
ihs5[ihs5$item_code==311 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7B'
ihs5[ihs5$item_code==411 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==407 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==408 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==301 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6B'
ihs5[ihs5$item_code==303 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7B'
ihs5[ihs5$item_code==413 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==601 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==901 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==313 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==102 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==104 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==406 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==608 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==307 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7B'
ihs5[ihs5$item_code==511 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==108 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==308 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7B'
ihs5[ihs5$item_code==402 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==103 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==202 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7'
ihs5[ihs5$item_code==312 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '7B'
ihs5[ihs5$item_code==606 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==801 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '6'


ihs5[ihs5$item_code==907 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unitA'] <- '15'
ihs5[ihs5$item_code==912 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unitA'] <- '15'
ihs5[ihs5$item_code==803 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unitA'] <- '15'
ihs5[ihs5$item_code==911 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unitA'] <- '15'
ihs5[ihs5$item_code==910 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unitA'] <- '15'
ihs5[ihs5$item_code==914 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unitA'] <- '15'
ihs5[ihs5$item_code==905 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unitA'] <- '15'
ihs5[ihs5$item_code==915 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unitA'] <- '15'
ihs5[ihs5$item_code==909 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unitA'] <- '15'
      ihs5$cons_unitA <- as.factor(ihs5$cons_unitA)
```


```{r}
ihs5[ihs5$cons_unit_oth == "PACKET", "cons_unit"] <- '51'
ihs5[ihs5$cons_unit_oth == "PACKET UNSPECIFIED", "cons_unit"] <- '51'
ihs5[ihs5$cons_unit_oth == "TABLESALT PACKET", "cons_unit"] <- '51'
ihs5[ihs5$cons_unit_oth == "PACKET MEDIUM", "cons_unit"] <- '51'
ihs5[ihs5$cons_unit_oth == "MEDIUM PACKET", "cons_unit"] <- '51'
ihs5[ihs5$cons_unit_oth == "SATCHET", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "TUB", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "TBE", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "SA HETS", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "TINA", "cons_unit"] <- '25'
ihs5[ihs5$cons_unit_oth == "GRAMS", "cons_unit"] <- '18'
ihs5[ihs5$cons_unit_oth == "G", "cons_unit"] <- '18'
ihs5[ihs5$cons_unit_oth == "PACKET (SMALL)", "cons_unit"] <- '54'
ihs5[ihs5$cons_unit_oth == "SMALL PACKET", "cons_unit"] <- '54'
ihs5[ihs5$cons_unit_oth == "SMALL PACKAGE", "cons_unit"] <- '54'
ihs5[ihs5$cons_unit_oth == "SMALL POCKET", "cons_unit"] <- '54'
ihs5[ihs5$cons_unit_oth == "SMALL  PACKET", "cons_unit"] <- '54'
ihs5[ihs5$cons_unit_oth == "PIECE", "cons_unit"] <- '9'
ihs5[ihs5$cons_unit_oth == "MANGO", "cons_unit"] <- '9'
ihs5[ihs5$cons_unit_oth == "BUNCH", "cons_unit"] <- '8'
ihs5[ihs5$cons_unit_oth == "BATCHES", "cons_unit"] <- '8'
ihs5[ihs5$cons_unit_oth == "BUNDLE", "cons_unit"] <- '8'
ihs5[ihs5$cons_unit_oth == "MILLILITRE", "cons_unit"] <- '19'
ihs5[ihs5$cons_unit_oth == "5L BUCKET", "cons_unit"] <- '26'
ihs5[ihs5$cons_unit_oth == "PAIL (SMALL)", "cons_unit"] <- '4'
ihs5[ihs5$cons_unit_oth == "TABLE SPOON", "cons_unit"] <- '59'
ihs5[ihs5$cons_unit_oth == "PAIL", "cons_unit"] <- '4'
ihs5[ihs5$cons_unit_oth == "KILOGRAM", "cons_unit"] <- '1'
ihs5[ihs5$cons_unit_oth == "PAIL (LARGE)", "cons_unit"] <- '4'
ihs5[ihs5$cons_unit_oth == "PAIL (MEDIUM)", "cons_unit"] <- '4'
ihs5[ihs5$cons_unit_oth == "LITRE", "cons_unit"] <- '15'
ihs5[ihs5$cons_unit_oth == "TINA FLAT", "cons_unit"] <- '25'
ihs5[ihs5$cons_unit_oth == "SMALL TINA FLAT", "cons_unit"] <- '25'
ihs5[ihs5$cons_unit_oth == "SATCHET (SMALL)", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "SMALL SACHET", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "SMALL SARCHET", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "TINA HEAPED", "cons_unit"] <- '25'
ihs5[ihs5$cons_unit_oth == "HEAP TINA", "cons_unit"] <- '25'
ihs5[ihs5$cons_unit_oth == "TINA (HEAPED)", "cons_unit"] <- '25'
ihs5[ihs5$cons_unit_oth == "CLUSTER", "cons_unit"] <- '44'
ihs5[ihs5$cons_unit_oth == "PACKET LARGE", "cons_unit"] <- '55'
ihs5[ihs5$cons_unit_oth == "BIG PACKET", "cons_unit"] <- '55'
ihs5[ihs5$cons_unit_oth == "SMALL HEAP", "cons_unit"] <- '10'
ihs5[ihs5$cons_unit_oth == "HEAP(SMALL)", "cons_unit"] <- '10'
ihs5[ihs5$cons_unit_oth == "LARGE HEAP", "cons_unit"] <- '10'
ihs5[ihs5$cons_unit_oth == "SMALL TIN", "cons_unit"] <- '71'
ihs5[ihs5$cons_unit_oth == "MEDIUM SACHET", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "MEDIAM SACHET", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "BUNCH SMALL", "cons_unit"] <- '8'
ihs5[ihs5$cons_unit_oth == "LARGE  SATCHET", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "SACHET  LARGE", "cons_unit"] <- '22'
ihs5[ihs5$cons_unit_oth == "HEAP (MEDIUM)", "cons_unit"] <- '10'
ihs5[ihs5$cons_unit_oth == "HEAP", "cons_unit"] <- '10'
ihs5[ihs5$cons_unit_oth == "50 GRAM", "cons_unit"] <- '42'

ihs5[ihs5$item_code==404 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==829 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==101 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==821 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==107 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==205 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==302 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==305 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==311 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==411 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==407 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==408 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==301 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==303 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==413 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==601 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==901 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==313 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==102 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==104 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==406 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==608 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==307 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==511 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==108 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==308 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==402 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==103 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==202 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==312 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '7'
ihs5[ihs5$item_code==606 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==801 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='PLATE', 'cons_unit'] <- '6'

ihs5[ihs5$item_code==907 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unit'] <- '15'
ihs5[ihs5$item_code==912 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unit'] <- '15'
ihs5[ihs5$item_code==803 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unit'] <- '15'
ihs5[ihs5$item_code==911 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unit'] <- '15'
ihs5[ihs5$item_code==910 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unit'] <- '15'
ihs5[ihs5$item_code==914 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unit'] <- '15'
ihs5[ihs5$item_code==905 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unit'] <- '15'
ihs5[ihs5$item_code==915 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unit'] <- '15'
ihs5[ihs5$item_code==909 & ihs5$cons_unit_oth=='BOTTLE (SMALL)','cons_unit'] <- '15'
      ihs5$cons_unitA <- as.factor(ihs5$cons_unitA)
         
```

>CLEANING ITEMS LABELED AS 'OTHER'

```{r}
item_other <- ihs5 %>% count(item_oth) %>% arrange(desc(n)) %>% print(n = nrow(ihs5))


```

```{r}

#Cleaning top 14 items labeled as 'Other'
ihs5$item_code[ihs5$item_oth=='SOYA'] <- 301
ihs5$item_code[ihs5$item_oth=='WILD LOQUAT'] <- 608
ihs5$item_code[ihs5$item_oth=='EGG PLANT'] <- 402
ihs5$item_code[ihs5$item_oth=='PEACHES'] <- 601
ihs5$item_code[ihs5$item_oth=='SORGHUM FLOUR'] <- 108
ihs5$item_code[ihs5$item_oth=='NKHUNGUZU'] <- 302
ihs5$item_code[ihs5$item_oth=='COW PEAS'] <- 308
ihs5$item_code[ihs5$item_oth=='FRIED MIXTURE OF MAIZE FLOUR/BANANA AND SODA'] <- 827
ihs5$item_code[ihs5$item_oth=='POWDERED JUICE'] <- 906
ihs5$item_code[ihs5$item_oth=='MUCUNA'] <- 837
ihs5$item_code[ihs5$item_oth=='IRISH POTATOES AND FLOUR'] <- 821
ihs5$item_code[ihs5$item_oth=='NKHWANI'] <- 404
ihs5$item_code[ihs5$item_oth=='CASSAVA LEAVES'] <- 404
ihs5$item_code[ihs5$item_oth=='FROZY'] <- 907


#Cleaning other items labeled as 'Other' 
ihs5$item_code[ihs5$item_oth=='GUAVAS'] <- 606
ihs5$item_code[ihs5$item_oth=='MIXTURE OF MAIZE FLOUR/BANANA/SUGAR'] <- 827
ihs5$item_code[ihs5$item_oth=='ZIBWENTE'] <- 821
ihs5$item_code[ihs5$item_oth=='ZIBHWENTE'] <- 821
ihs5$item_code[ihs5$item_oth=='ZIBWEMPWE'] <- 821
ihs5$item_code[ihs5$item_oth=='ZIGEDE'] <- 821
ihs5$item_code[ihs5$item_oth=='CHIGEGE'] <- 821
ihs5$item_code[ihs5$item_oth=='WINE'] <- 914
ihs5$item_code[ihs5$item_oth=='GROUND AND COOKED BEANS'] <- 302
ihs5$item_code[ihs5$item_oth=='KHOBWE'] <- 308
ihs5$item_code[ihs5$item_oth=='SOHGUM FLOUR (UFA)'] <- 108
ihs5$item_code[ihs5$item_oth=='NTAPASHA (CASSAVA LEAVES)'] <- 404
ihs5$item_code[ihs5$item_oth=='MKHAKA'] <- 409
ihs5$item_code[ihs5$item_oth=='NTAPASHA'] <- 404
ihs5$item_code[ihs5$item_oth=='RAPE'] <- 403
ihs5$item_code[ihs5$item_oth=='NANAZI'] <- 604
ihs5$item_code[ihs5$item_oth=='PEAGION PEAS'] <- 303
ihs5$item_code[ihs5$item_oth=='MAPIRA'] <- 108
ihs5$item_code[ihs5$item_oth=='GRASSHOPPER'] <- 511
ihs5$item_code[ihs5$item_oth=='LOCUST'] <- 511
ihs5$item_code[ihs5$item_oth=='ZIWALA'] <- 511
ihs5$item_code[ihs5$item_oth=='MPILU'] <- 403
ihs5$item_code[ihs5$item_oth=='KHUNGUZU'] <- 302
ihs5$item_code[ihs5$item_oth=='MISALE'] <- 108
ihs5$item_code[ihs5$item_oth=='KHOLOWA'] <- 404
ihs5$item_code[ihs5$item_oth=='CHISOSO'] <- 404
ihs5$item_code[ihs5$item_oth=='BONONGWE'] <- 404
ihs5$item_code[ihs5$item_oth=='KHWANYA'] <- 404
ihs5$item_code[ihs5$item_oth=='KWANYA'] <- 404
ihs5$item_code[ihs5$item_oth=='MTAMBE'] <- 404
ihs5$item_code[ihs5$item_oth=='LUNI'] <- 404
ihs5$item_code[ihs5$item_oth=='MPHONDA'] <- 410
ihs5$item_code[ihs5$item_oth=='EGGS PLANTS'] <- 402
ihs5$item_code[ihs5$item_oth=='THERERE'] <- 411
ihs5$item_code[ihs5$item_oth=='CHAMWAMBA'] <- 411
ihs5$item_code[ihs5$item_oth=='CHITAMBE'] <- 404
ihs5$item_code[ihs5$item_oth=='DORGHUM FLOUR'] <- 108
ihs5$item_code[ihs5$item_oth=='GATHERED THERERE'] <- 411
ihs5$item_code[ihs5$item_oth=='KATATA'] <- 411
ihs5$item_code[ihs5$item_oth=='MABILINGANO'] <- 402
ihs5$item_code[ihs5$item_oth=='PUMPKINS'] <- 410
ihs5$item_code[ihs5$item_oth=='THERERE WACHINYOLOMONYA'] <- 411
ihs5$item_code[ihs5$item_oth=='THERERE WAMASAMBA'] <- 411
ihs5$item_code[ihs5$item_oth=='WILD THERERE'] <- 411
ihs5$item_code[ihs5$item_oth=='DUCK'] <- 509
ihs5$item_code[ihs5$item_oth=='ANTELOPE'] <- 505
ihs5$item_code[ihs5$item_oth=='BIRD'] <- 509
ihs5$item_code[ihs5$item_oth=='TSETSENYA (LOOKS LIKE IMSECTS'] <- 511
ihs5$item_code[ihs5$item_oth=='CHIBUKU BEER'] <- 908
ihs5$item_code[ihs5$item_oth=='FIRST CHOICE MILK'] <- 701
ihs5$item_code[ihs5$item_oth=='SODA'] <- 812
ihs5$item_code[ihs5$item_oth=='ROYCO'] <- 811
ihs5$item_code[ihs5$item_oth=='MIXED BEANS'] <- 302
ihs5$item_code[ihs5$item_oth=='BUTTER BEANS'] <- 301
ihs5$item_code[ihs5$item_oth=='RED BEANS'] <- 302
ihs5$item_code[ihs5$item_oth=='YELLOW BEANS'] <- 302
ihs5$item_code[ihs5$item_oth=='ROASTED SWEET POTATO'] <- 832
ihs5$item_code[ihs5$item_oth=='IRISH POTATOES'] <- 205
ihs5$item_code[ihs5$item_oth=='PEPPER'] <- 811
ihs5$item_code[ihs5$item_oth=='YAMS'] <- 208
ihs5$item_code[ihs5$item_oth=='COCOYAMS'] <- 208
ihs5$item_code[ihs5$item_oth=='COOKED SWEET POTATOES (VENDORS)'] <- 831
ihs5$item_code[ihs5$item_oth=='NKHWALI'] <- 509
ihs5$item_code[ihs5$item_oth=='BAKES'] <- 906
ihs5$item_code[ihs5$item_oth=='BAKES JUICE'] <- 906
ihs5$item_code[ihs5$item_oth=='DRAGON  FROZY'] <- 907
ihs5$item_code[ihs5$item_oth=='JOLKY JUS'] <- 906
ihs5$item_code[ihs5$item_oth=='JOLLY JAS JUICE'] <- 906
ihs5$item_code[ihs5$item_oth=='THUMPS UP'] <- 907
ihs5$item_code[ihs5$item_oth=='WAKA JUICE'] <- 906
ihs5$item_code[ihs5$item_oth=='MEXICAN APPLE'] <- 609
ihs5$item_code[ihs5$item_oth=='CUSTARD APPLE'] <- 609
ihs5$item_code[ihs5$item_oth=='LEMON'] <- 603
ihs5$item_code[ihs5$item_oth=='WILD FRUIT'] <- 608
ihs5$item_code[ihs5$item_oth=='MAXICAN APPLE'] <- 609
ihs5$item_code[ihs5$item_oth=='WILD PASSION FRUIT'] <- 608
ihs5$item_code[ihs5$item_oth=='MEAN APPLE'] <- 609
ihs5$item_code[ihs5$item_oth=='CUCUMBER'] <- 409
ihs5$item_code[ihs5$item_oth=='MTAMBE'] <- 404
ihs5$item_code[ihs5$item_oth=='PINEAPPLE'] <- 604

```

>DONE : ARCHIVE 

```{r}

write.csv(ihs5, here::here('hh_mod_g_initial.csv'))

```

