---
title: "ameIHS5"
author: "Gareth Osman"
date: "8/5/2021"
output: html_document
---

> Packages

```{r}
library(epiDisplay)
library(foreign)
library(readxl)
library(psych)
library(tidyverse)
```

>Data 

```{r}

ihs5.roster <- haven::read_dta(here::here('HH_MOD_B.dta'))

ihs5_health <-haven::read_dta(here::here('HH_MOD_D.dta'))

ame <- read.csv(here::here('ihs5.ame.factors.csv'))

ame.spec <- read.csv(here::here('ihs5.ame.spec.csv'))
names(ame.spec) [1] <- 'cat'
ame.spec <- ame.spec %>% select(cat, ame.spec, afe.spec)


```


> HH roster data

```{r}
ihs5.roster$case_id <- as.character(ihs5.roster$case_id)

#Sex
names(ihs5.roster)[names(ihs5.roster) == 'hh_b03'] <- 'sex'

#Age
names(ihs5.roster)[names(ihs5.roster) == 'hh_b05a'] <- 'age_y'
names(ihs5.roster)[names(ihs5.roster) == 'hh_b05b'] <- 'age_m'
ihs5.roster.c <- ihs5.roster %>% select(case_id, HHID, sex, age_y, age_m)

ihs5.roster.c$age_y <- as.numeric(ihs5.roster.c$age_y)
ihs5.roster.c$age_m <- as.numeric(ihs5.roster.c$age_m)
ihs5.roster.c <- ihs5.roster.c %>% mutate(age.m.total = (age_y*12 + age_m))
ihs5.roster.c <- ihs5.roster.c %>% mutate(age.u2 = case_when(age_y < 2 ~ "TRUE"))
#Per capita variable
ihs5.roster.c <- ihs5.roster.c %>% mutate(pc = 1)
```

> Per capita variable

```{r}
hh.pc <- aggregate(ihs5.roster.c$pc, by=list(HHID=ihs5.roster.c$HHID), FUN=sum)

names(hh.pc)[names(hh.pc) == 'sum.ihs5.roster.c$pc'] <- 'pc'
```

>Merge HH demographic data with AME/AFE factors

Men's weight: 65kg (assumption)
Women's weight: 55kg (from DHS)
PAL: 1.6X the BMR
```{r}
ihs5.roster.c$age_y <- as.factor(ihs5.roster.c$age_y)
names(ame) [1] <- 'age'
ame$age <- as.factor(ame$age)

ihs5.roster.c <- merge(x=ihs5.roster.c , y=ame , by.x='age_y', by.y='age', fill=-9999, all.x = TRUE) %>% arrange(case_id) %>% select(case_id, HHID, sex, age_y, age_m, age.m.total, age.u2, ame.m, ame.f, afe.m, afe.f)

ihs5.roster.c <- ihs5.roster.c  %>%
  mutate(ame.base = ifelse(sex == 1, ame.m,
               ifelse(sex == 2, ame.f, NA)))

ihs5.roster.c <- ihs5.roster.c  %>%
  mutate(afe.base = ifelse(sex == 1, afe.m,
               ifelse(sex == 2, afe.f, NA)))
```

> Dietary energy requirements for children under 1 

```{r}
#create variable for under 1 year old categories
ihs5.roster.c <- ihs5.roster.c %>% mutate(age.u1.cat = case_when(age.m.total  < 6 ~ "0-5 months",
                                          age.m.total  >= 6 & age.m.total  < 9 ~ "6-8 months",
                                          age.m.total  >= 9 & age.m.total  < 12 ~ "9-11 months")) 

ihs5.roster.c <- merge(x=ihs5.roster.c , y=ame.spec , by.x='age.u1.cat', by.y='cat', fill=-9999, all.x = TRUE) %>% arrange(case_id)
```

> Extra energy required for lactation

```{r}
ihs5.roster.c <- ihs5.roster.c %>% mutate(ame.lac = case_when(age.u2 =="TRUE" ~ 0.19))
ihs5.roster.c <- ihs5.roster.c %>% mutate(afe.lac = case_when(age.u2 =="TRUE" ~ 0.24))
```

> Clean and aggregate AME/AFE values

```{r}
ihs5.roster.c$ame.spec[is.na(ihs5.roster.c$ame.spec)] <- 0
ihs5.roster.c$afe.spec[is.na(ihs5.roster.c$afe.spec)] <- 0
ihs5.roster.c$ame.lac[is.na(ihs5.roster.c$ame.lac)] <- 0
ihs5.roster.c$afe.lac[is.na(ihs5.roster.c$afe.lac)] <- 0

ihs5.roster.c <- ihs5.roster.c %>% mutate(ame = ame.base + ame.spec + ame.lac)
ihs5.roster.c <- ihs5.roster.c %>% mutate(afe = afe.base + afe.spec + afe.lac)

hh.ame <- aggregate(ihs5.roster.c$ame, by=list(HHID=ihs5.roster.c$HHID), FUN=sum)
hh.afe <- aggregate(ihs5.roster.c$afe, by=list(HHID=ihs5.roster.c$HHID), FUN=sum)
names(hh.ame)[names(hh.ame) == 'sum.ihs5.roster.c$ame'] <- 'ame.x'
names(hh.afe)[names(hh.afe) == 'sum.ihs5.roster.c$afe'] <- 'afe.x'
hme <- merge(x=hh.ame , y=hh.afe  , by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
```


>Extra energy requirements for pregnancy

```{r}
ihs5_health$case_id <- as.character(ihs5_health$case_id)

#Illness
names(ihs5_health)[names(ihs5_health) == 'hh_d05a'] <- 'ill1'
names(ihs5_health)[names(ihs5_health) == 'hh_d05b'] <- 'ill2'
ihs5_health.c <- ihs5_health %>% select(case_id, HHID, ill1, ill2, hh_d05_oth)
ihs5.preg <- ihs5_health.c %>% filter(ill1==28 | ill2==28)

ihs5.preg$ame.preg <- 0.11
ihs5.preg$afe.preg <- 0.14
ihs5.preg <- ihs5.preg %>% select(HHID, ame.preg, afe.preg)

hme <- merge(x=hme , y=ihs5.preg , by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE) 
hme$ame.preg[is.na(hme$ame.preg)] <- 0
hme$afe.preg[is.na(hme$afe.preg)] <- 0

hme$ame <- hme$ame.x + hme$ame.preg
hme$afe <- hme$afe.x + hme$afe.preg

```

>Finalization

```{r}
hme <- hme %>% select(HHID, ame, afe)
hme <- merge(x=hme , y=hh.pc , by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
```

> Fix single household factors

```{r}
hme <- hme %>% mutate(ame = case_when(pc == 1 ~ 1, pc != 1 ~ ame))
hme <- hme %>% mutate(afe = case_when(pc == 1 ~ 1, pc != 1 ~ afe))
```

> DONE: archive

```{r}
write.csv(hme, here::here('hme.final.csv'))

```
