---
title: "ihs5_cleaning_purchased"
author: "Gareth Osman"
date: "5/21/2021"
output: html_document
---


INTRO

We need to clean separately how much items came from purchases to be able to analyse their contribution to the total household intake. 

```{r}
#--- Load libraries 
library(plyr)
library(haven)
library(magrittr)
library(foreign)
library(psych)
library(readxl)
library(labelled)
library(tidyverse)

#--- Change to show no scientific notation & round to 3 decimal places
options(scipen = 10, digits=3) 

#--- Read in the files
ihs5 <- haven::read_dta("C:/Users/Gareth Osman/OneDrive - LUANAR/Documents/MAPS Project/ihs5 Cleaning/HH_MOD_G1.dta")
 hhfilt <- haven::read_dta("C:/Users/Gareth Osman/OneDrive - LUANAR/Documents/MAPS Project/ihs5 Cleaning/hh_mod_a_filt.dta")
convf <- haven::read_dta("C:/Users/Gareth Osman/OneDrive - LUANAR/Documents/MAPS Project/ihs5 Cleaning/ihs_foodconversion_factor_2020.dta")

```

>CHANGE VARIBLE NAMES AND DATA CLASS

```{r}
#--- Renaming variables to standard names 

#Summary info 
 ihs5 <- ihs5 %>% rename(
      id = 'case_id',
      consYN = 'hh_g01',
      item_code = 'hh_g02',
      item_oth = 'hh_g01_oth')


ihs5$id <- as.character(ihs5$id)
ihs5$consYN <- as.factor(ihs5$consYN)
ihs5$item_name <- as.factor(ihs5$item_code)
ihs5$item_oth <- as.factor(ihs5$item_oth)


#G03: Quantity and units consumed of each food 
ihs5 <- ihs5 %>% rename(
      cons_quant = 'hh_g03a',
      cons_unitA = 'hh_g03b',
      cons_unit = 'hh_g03b_label',
      cons_unit_oth = 'hh_g03b_oth',
      cons_unit_size = 'hh_g03c')

ihs5$cons_quant <- as.numeric(ihs5$cons_quant)
ihs5$cons_unitA <- as.factor(ihs5$cons_unitA)
ihs5$cons_unit <- as.factor(ihs5$cons_unit)
ihs5$cons_unit_oth <- as.factor(ihs5$cons_unit_oth)
ihs5$cons_unit_size <- as.factor(ihs5$cons_unit_size)


#G04: Quantity and units purchased of each food 
ihs5 <- ihs5 %>% rename(
      purc_quant = 'hh_g04a',
      purc_unitA = 'hh_g04b',
      purc_unit = 'hh_g04b_label',
      purc_unit_oth = 'hh_g04b_oth',
      purc_unit_size = 'hh_g04c',
      purc_price = 'hh_g05')

ihs5$purc_quant <- as.numeric(ihs5$purc_quant)
ihs5$purc_unitA <- as.factor(ihs5$purc_unitA)
ihs5$purc_unit <- as.factor(ihs5$purc_unit)
ihs5$purc_unit_oth <- as.factor(ihs5$purc_unit_oth)
ihs5$purc_unit_size <- as.factor(ihs5$purc_unit_size)
ihs5$purc_price <- as.numeric(ihs5$purc_price)

#G06: Quantity and units home produced of each food
ihs5 <- ihs5 %>% rename(
      prod_quant = 'hh_g06a',
      prod_unitA = 'hh_g06b',
      prod_unit = 'hh_g06b_label',
      prod_unit_oth = 'hh_g06b_oth',
      prod_unit_size = 'hh_g06c')

ihs5$prod_quant <- as.numeric(ihs5$prod_quant)
ihs5$prod_unitA <- as.factor(ihs5$prod_unitA)
ihs5$prod_unit <- as.factor(ihs5$prod_unit)
ihs5$prod_unit_oth <-as.factor(ihs5$prod_unit_oth)
ihs5$prod_unit_size <- as.factor(ihs5$prod_unit_size)

#G07: Quantity and units gifted/other source of each food 
ihs5 <- ihs5 %>% rename(
      gift_quant = 'hh_g07a',
      gift_unitA = 'hh_g07b',
      gift_unit = 'hh_g07b_label',
      gift_unit_oth = 'hh_g07b_oth',
      gift_unit_size = 'hh_g07c')

ihs5$gift_quant <- as.numeric(ihs5$gift_quant)
ihs5$gift_unitA <- as.factor(ihs5$gift_unitA)
ihs5$gift_unit <- as.factor(ihs5$gift_unit)
ihs5$gift_unit_oth <- as.factor(ihs5$gift_unit_oth)
ihs5$gift_unit_size <- as.factor(ihs5$gift_unit_size)

#Removing a duplicate variable- item_code
 ihs5 <- select(ihs5, -33)

```

>EXTRACTING VALUES AND LABELS OF LABELED 'ITEM_CODE'

The ihs5 dataset has stored the variable 'item_code' as a label (dbl+lbl). Hence we need to extract a two column data frame of value and label. The label and values are stored as an attribute of data frame. 

```{r}
ihs5 %>% distinct(item_code)

ihs5_attr <- stack(attr(ihs5$item_code, 'labels')) 

#Changing variable class
ihs5_attr$values <- as.factor(ihs5_attr$values)
ihs5$item_code <- as.factor(ihs5$item_code)

#Merging the values and labels back into the ihs5 dataset
ihs5 %>% 
    left_join(., ihs5_attr, by=c("item_code" = "values"))
ihs5 <- ihs5 %>% 
     left_join(., ihs5_attr, by=c("item_code" = "values"))

ihs5 <- ihs5 %>% relocate("ind", .after = item_code)
ihs5 <- ihs5 %>% rename(item_name_original = 'ind')

ihs5 <- ihs5 %>% arrange(cons_unit) %>% arrange(item_code) %>% arrange(HHID)
ihs5 <- ihs5 %>% select(id:consYN, item_code, item_name_original, item_oth:cons_unitA, cons_unit, cons_unit_oth:gift_unit_size)

```

 >CLEANING UNITS LABELED AS OTHER - VARIABLE 'PURCHASED'

```{r}
ihs5 %>% filter(purc_unitA==23)

ihs5 %>% filter(purc_unit_oth!= "") %>% count(purc_unit_oth) %>% arrange(desc(n)) %>% print(n = nrow(ihs5))

#units listed as "other" and have not been converted to standard units
ihs5 %>% filter(purc_unitA==23) %>% count(purc_unit_oth) %>% arrange(desc(n)) %>% print(n = nrow(ihs5))
 
ihs5 %>% filter(., purc_unit_oth=="HEAP") %>% select(item_name_original, item_code, purc_quant, purc_unitA, purc_unit, purc_unit_oth) %>% arrange(item_name_original) %>% print(n = nrow(ihs5))

ihs5 %>% filter(item_code=="803") %>% select(item_name_original, item_code, purc_quant, purc_unitA, purc_unit, purc_unit_oth) %>% arrange(desc(purc_unitA))  %>% print(n = nrow(ihs5))
 
```

```{r}

#Cleaning top 13 known units labeled as 'Other'
ihs5[ihs5$purc_unit_oth == "PACKET", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "SATCHET", "purc_unitA"] <- "22"
ihs5[ihs5$purc_unit_oth == "PACKET UNSPECIFIED", "purc_unitA"] <- "51"
ihs5[ihs5$purc_unit_oth == "GRAMS", "purc_unitA"] <- "18"
ihs5[ihs5$purc_unit_oth == "PACKET (SMALL)", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "TINA", "purc_unitA"] <- "25"
ihs5[ihs5$purc_unit_oth == "PIECE", "purc_unitA"] <- "9"
ihs5[ihs5$purc_unit_oth == "MILLILITRE", "purc_unitA"] <- "19"
ihs5[ihs5$purc_unit_oth == "PLATE", "purc_unitA"] <- "6B"
ihs5[ihs5$purc_unit_oth == "BASIN", "purc_unitA"] <- "27D"
ihs5[ihs5$purc_unit_oth == "HEAP", "purc_unitA"] <- "10"
ihs5[ihs5$purc_unit_oth == "BOTTLE (SMALL)" & ihs5$item_code == "803", "purc_unitA"] <- "15"

#Bunch- 8 doesn't exist in the standard unit list, but it does in the conversion factor file. Hence we store it.
ihs5$purc_unitA <- as.character(ihs5$purc_unitA) 
ihs5[ihs5$purc_unit_oth == "BUNCH", "purc_unitA"] <- "8"
ihs5$purc_unitA <- as.factor(ihs5$purc_unitA)


#Cleaning other known units labeled as 'Other'
ihs5[ihs5$purc_unit_oth == "TABLE SPOON", "purc_unitA"] <- "59"
ihs5[ihs5$purc_unit_oth == "SMALL PACKET", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "5L BUCKET", "purc_unitA"] <- "26"
ihs5[ihs5$purc_unit_oth == "PAIL (SMALL)", "purc_unitA"] <- "4A"
ihs5[ihs5$purc_unit_oth == "LITRE", "purc_unitA"] <- "15"
ihs5[ihs5$purc_unit_oth == "SATCHET (SMALL)", "purc_unitA"] <- "22A"
ihs5[ihs5$purc_unit_oth == "TINA FLAT", "purc_unitA"] <- "25A"
ihs5[ihs5$purc_unit_oth == "KILOGRAM", "purc_unitA"] <- "1"
ihs5[ihs5$purc_unit_oth == "TINA HEAPED", "purc_unitA"] <- "25B"
ihs5[ihs5$purc_unit_oth == "BATCHES", "purc_unitA"] <- "8"
ihs5[ihs5$purc_unit_oth == "PAIL (MEDIUM)", "purc_unitA"] <- "4B"
ihs5[ihs5$purc_unit_oth == "SMALL SACHET", "purc_unitA"] <- "22A"
ihs5[ihs5$purc_unit_oth == "BUNDLE", "purc_unitA"] <- "8"
ihs5[ihs5$purc_unit_oth == "PAIL", "purc_unitA"] <- "4"
ihs5[ihs5$purc_unit_oth == "MANGO", "purc_unitA"] <- "9"
ihs5[ihs5$purc_unit_oth == "SMALL  PACKET", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "SMALL HEAP", "purc_unitA"] <- "10A"
ihs5[ihs5$purc_unit_oth == "SMALL TIN", "purc_unitA"] <- "71"
ihs5[ihs5$purc_unit_oth == "HEAP (SMALL)", "purc_unitA"] <- "10A"
ihs5[ihs5$purc_unit_oth == "CLUSTER", "purc_unitA"] <- "44"
ihs5[ihs5$purc_unit_oth == "MEDIUM SACHET", "purc_unitA"] <- "22B"
ihs5[ihs5$purc_unit_oth == "PACKET LARGE", "purc_unitA"] <- "55"
ihs5[ihs5$purc_unit_oth == "PAIL (LARGE)", "purc_unitA"] <- "4C"
ihs5[ihs5$purc_unit_oth == "SMALL SARCHET", "purc_unitA"] <- "22A"
ihs5[ihs5$purc_unit_oth == "25 GRAM SATCHET", "purc_unitA"] <- "41"
ihs5[ihs5$purc_unit_oth == "BANCHES", "purc_unitA"] <- "8"
ihs5[ihs5$purc_unit_oth == "BIG PACKET", "purc_unitA"] <- "55"
ihs5[ihs5$purc_unit_oth == "BRUNCH", "purc_unitA"] <- "8"
ihs5[ihs5$purc_unit_oth == "G", "purc_unitA"] <- "18"
ihs5[ihs5$purc_unit_oth == "HEAP'S", "purc_unitA"] <- "10"
ihs5[ihs5$purc_unit_oth == "HEAP (MEDIUM)", "purc_unitA"] <- "10B"
ihs5[ihs5$purc_unit_oth == "HEAP(SMALL)", "purc_unitA"] <- "10A"
ihs5[ihs5$purc_unit_oth == "KKILOGRAM", "purc_unitA"] <- "1"
ihs5[ihs5$purc_unit_oth == "LARGE SATCHET", "purc_unitA"] <- "22C"
ihs5[ihs5$purc_unit_oth == "LITRE'S", "purc_unitA"] <- "15"
ihs5[ihs5$purc_unit_oth == "MEDIAM SACHET", "purc_unitA"] <- "22B"
ihs5[ihs5$purc_unit_oth == "PACCKET", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACHE", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKERS", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKERTS", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKEYS", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKKET", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PATCKET", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PCKETT", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "POCKET", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PWCKETS", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKET TABLE SALT", "purc_unitA"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKET SMAlL", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "SACHER", "purc_unitA"] <- "22"
ihs5[ihs5$purc_unit_oth == "SACHEY", "purc_unitA"] <- "22"
ihs5[ihs5$purc_unit_oth == "SACHET  LARGE", "purc_unitA"] <- "22C"
ihs5[ihs5$purc_unit_oth == "SMALL PACKET GOLD TEA", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "SMALL TABLE SALT PACKET", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "SMALL PAIL.", "purc_unitA"] <- "4A"
ihs5[ihs5$purc_unit_oth == "TIBE", "purc_unitA"] <- "22"
ihs5[ihs5$purc_unit_oth == "TUB", "purc_unitA"] <- "22"


#Cleaning unknown units labeled as 'Other' with proxy units 
ihs5[ihs5$purc_unit_oth == "MEDIUM TINS", "purc_unitA"] <- "72"
ihs5[ihs5$purc_unit_oth == "PACKET MEDIUM", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "MEDIUM PACKET", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "PACKET (MEDIUM)", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "PACKET MDIUM", "purc_unitA"] <- "54"
ihs5[ihs5$purc_unit_oth == "125G PACKET", "purc_unitA"] <- "34"
ihs5[ihs5$purc_unit_oth == "HALF LOAF", "purc_unitA"] <- "31"
ihs5[ihs5$purc_unit_oth == "10G PACKET", "purc_unitA"] <- "70"
ihs5[ihs5$purc_unit_oth == "80G PACKET", "purc_unitA"] <- "34"
ihs5[ihs5$purc_unit_oth == "90 ZZGRAM PACKET", "purc_unitA"] <- "34"
ihs5[ihs5$purc_unit_oth == "PACKET OF 125 GRAMS", "purc_unitA"] <- "34"
ihs5[ihs5$purc_unit_oth == "HALF PAIL", "purc_unitA"] <- "4B"
ihs5[ihs5$purc_unit_oth == "SMALL TIIN", "purc_unitA"] <- "71"
ihs5[ihs5$purc_unit_oth == "SMALL TINA FLAT", "purc_unitA"] <- "25A"

```

```{r}
#Cleaning top 13 known units labeled as 'Other'
ihs5[ihs5$purc_unit_oth == "PACKET", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "SATCHET", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "PACKET UNSPECIFIED", "purc_unit"] <- "51"
ihs5[ihs5$purc_unit_oth == "GRAMS", "purc_unit"] <- "18"
ihs5[ihs5$purc_unit_oth == "PACKET (SMALL)", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "TINA", "purc_unit"] <- "25"
ihs5[ihs5$purc_unit_oth == "PIECE", "purc_unit"] <- "9"
ihs5[ihs5$purc_unit_oth == "MILLILITRE", "purc_unit"] <- "19"
ihs5[ihs5$purc_unit_oth == "PLATE", "purc_unit"] <- "6"
ihs5[ihs5$purc_unit_oth == "BASIN", "purc_unit"] <- "27"
ihs5[ihs5$purc_unit_oth == "HEAP", "purc_unit"] <- "10"
ihs5[ihs5$purc_unit_oth == "BUNCH", "purc_unit"] <- "8"
ihs5[ihs5$purc_unit_oth == "BOTTLE (SMALL)" & ihs5$item_code == "803", "purc_unit"] <- "15"


#Cleaning other known units labeled as 'Other'
ihs5[ihs5$purc_unit_oth == "TABLE SPOON", "purc_unit"] <- "59"
ihs5[ihs5$purc_unit_oth == "SMALL PACKET", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "5L BUCKET", "purc_unit"] <- "26"
ihs5[ihs5$purc_unit_oth == "PAIL (SMALL)", "purc_unit"] <- "4"
ihs5[ihs5$purc_unit_oth == "LITRE", "purc_unit"] <- "15"
ihs5[ihs5$purc_unit_oth == "SATCHET (SMALL)", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "TINA FLAT", "purc_unit"] <- "25"
ihs5[ihs5$purc_unit_oth == "KILOGRAM", "purc_unit"] <- "1"
ihs5[ihs5$purc_unit_oth == "TINA HEAPED", "purc_unit"] <- "25"
ihs5[ihs5$purc_unit_oth == "BATCHES", "purc_unit"] <- "8"
ihs5[ihs5$purc_unit_oth == "PAIL (MEDIUM)", "purc_unit"] <- "4"
ihs5[ihs5$purc_unit_oth == "SMALL SACHET", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "BUNDLE", "purc_unit"] <- "8"
ihs5[ihs5$purc_unit_oth == "PAIL", "purc_unit"] <- "4"
ihs5[ihs5$purc_unit_oth == "MANGO", "purc_unit"] <- "9"
ihs5[ihs5$purc_unit_oth == "SMALL  PACKET", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "SMALL HEAP", "purc_unit"] <- "10"
ihs5[ihs5$purc_unit_oth == "SMALL TIN", "purc_unit"] <- "71"
ihs5[ihs5$purc_unit_oth == "HEAP (SMALL)", "purc_unit"] <- "10"
ihs5[ihs5$purc_unit_oth == "CLUSTER", "purc_unit"] <- "44"
ihs5[ihs5$purc_unit_oth == "MEDIUM SACHET", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "PACKET LARGE", "purc_unit"] <- "55"
ihs5[ihs5$purc_unit_oth == "PAIL (LARGE)", "purc_unit"] <- "4"
ihs5[ihs5$purc_unit_oth == "SMALL SARCHET", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "25 GRAM SATCHET", "purc_unit"] <- "41"
ihs5[ihs5$purc_unit_oth == "BANCHES", "purc_unit"] <- "8"
ihs5[ihs5$purc_unit_oth == "BIG PACKET", "purc_unit"] <- "55"
ihs5[ihs5$purc_unit_oth == "BRUNCH", "purc_unit"] <- "8"
ihs5[ihs5$purc_unit_oth == "G", "purc_unit"] <- "18"
ihs5[ihs5$purc_unit_oth == "HEAP'S", "purc_unit"] <- "10"
ihs5[ihs5$purc_unit_oth == "HEAP (MEDIUM)", "purc_unit"] <- "10"
ihs5[ihs5$purc_unit_oth == "HEAP(SMALL)", "purc_unit"] <- "10"
ihs5[ihs5$purc_unit_oth == "KKILOGRAM", "purc_unit"] <- "1"
ihs5[ihs5$purc_unit_oth == "LARGE SATCHET", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "LITRE'S", "purc_unit"] <- "15"
ihs5[ihs5$purc_unit_oth == "MEDIAM SACHET", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "PACCKET", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACHE", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKERS", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKERTS", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKEYS", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKKET", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PATCKET", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PCKETT", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "POCKET", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PWCKETS", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKET TABLE SALT", "purc_unit"] <- "60"
ihs5[ihs5$purc_unit_oth == "PACKET SMAlL", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "SACHER", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "SACHEY", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "SACHET  LARGE", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "SMALL PACKET GOLD TEA", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "SMALL TABLE SALT PACKET", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "SMALL PAIL.", "purc_unit"] <- "4"
ihs5[ihs5$purc_unit_oth == "TIBE", "purc_unit"] <- "22"
ihs5[ihs5$purc_unit_oth == "TUB", "purc_unit"] <- "22"


#Cleaning unknown units labeled as 'Other' with proxy units 
ihs5[ihs5$purc_unit_oth == "MEDIUM TINS", "purc_unit"] <- "72"
ihs5[ihs5$purc_unit_oth == "PACKET MEDIUM", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "MEDIUM PACKET", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "PACKET (MEDIUM)", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "PACKET MDIUM", "purc_unit"] <- "54"
ihs5[ihs5$purc_unit_oth == "125G PACKET", "purc_unit"] <- "34"
ihs5[ihs5$purc_unit_oth == "HALF LOAF", "purc_unit"] <- "31"
ihs5[ihs5$purc_unit_oth == "10G PACKET", "purc_unit"] <- "70"
ihs5[ihs5$purc_unit_oth == "80G PACKET", "purc_unit"] <- "34"
ihs5[ihs5$purc_unit_oth == "90 ZZGRAM PACKET", "purc_unit"] <- "34"
ihs5[ihs5$purc_unit_oth == "PACKET OF 125 GRAMS", "purc_unit"] <- "34"
ihs5[ihs5$purc_unit_oth == "HALF PAIL", "purc_unit"] <- "4"
ihs5[ihs5$purc_unit_oth == "SMALL TIIN", "purc_unit"] <- "71"
ihs5[ihs5$purc_unit_oth == "SMALL TINA FLAT", "purc_unit"] <- "25"

```

>CLEANING ITEMS LABELED AS 'OTHER'

```{r}
item_other <- ihs5 %>% count(item_oth) %>% arrange(desc(n)) %>% print(n = nrow(ihs5))

```


```{r}

#Cleaning top 14 items labeled as 'Other'
ihs5$item_code[ihs5$item_oth=='SOYA'] <- 301
ihs5$item_code[ihs5$item_oth=='WILD LOQUAT'] <- 608
ihs5$item_code[ihs5$item_oth=='EGG PLANT'] <- 402
ihs5$item_code[ihs5$item_oth=='PEACHES'] <- 601
ihs5$item_code[ihs5$item_oth=='SORGHUM FLOUR'] <- 108
ihs5$item_code[ihs5$item_oth=='NKHUNGUZU'] <- 302
ihs5$item_code[ihs5$item_oth=='COW PEAS'] <- 308
ihs5$item_code[ihs5$item_oth=='FRIED MIXTURE OF MAIZE FLOUR/BANANA AND SODA'] <- 827
ihs5$item_code[ihs5$item_oth=='POWDERED JUICE'] <- 906
ihs5$item_code[ihs5$item_oth=='MUCUNA'] <- 837
ihs5$item_code[ihs5$item_oth=='IRISH POTATOES AND FLOUR'] <- 821
ihs5$item_code[ihs5$item_oth=='NKHWANI'] <- 404
ihs5$item_code[ihs5$item_oth=='CASSAVA LEAVES'] <- 404
ihs5$item_code[ihs5$item_oth=='FROZY'] <- 907


#Cleaning other items labeled as 'Other' 
ihs5$item_code[ihs5$item_oth=='GUAVAS'] <- 606
ihs5$item_code[ihs5$item_oth=='MIXTURE OF MAIZE FLOUR/BANANA/SUGAR'] <- 827
ihs5$item_code[ihs5$item_oth=='ZIBWENTE'] <- 821
ihs5$item_code[ihs5$item_oth=='ZIBHWENTE'] <- 821
ihs5$item_code[ihs5$item_oth=='ZIBWEMPWE'] <- 821
ihs5$item_code[ihs5$item_oth=='ZIGEDE'] <- 821
ihs5$item_code[ihs5$item_oth=='CHIGEGE'] <- 821
ihs5$item_code[ihs5$item_oth=='WINE'] <- 914
ihs5$item_code[ihs5$item_oth=='GROUND AND COOKED BEANS'] <- 302
ihs5$item_code[ihs5$item_oth=='KHOBWE'] <- 308
ihs5$item_code[ihs5$item_oth=='SOHGUM FLOUR (UFA)'] <- 108
ihs5$item_code[ihs5$item_oth=='NTAPASHA (CASSAVA LEAVES)'] <- 404
ihs5$item_code[ihs5$item_oth=='MKHAKA'] <- 409
ihs5$item_code[ihs5$item_oth=='NTAPASHA'] <- 404
ihs5$item_code[ihs5$item_oth=='RAPE'] <- 403
ihs5$item_code[ihs5$item_oth=='NANAZI'] <- 604
ihs5$item_code[ihs5$item_oth=='PEAGION PEAS'] <- 303
ihs5$item_code[ihs5$item_oth=='MAPIRA'] <- 108
ihs5$item_code[ihs5$item_oth=='GRASSHOPPER'] <- 511
ihs5$item_code[ihs5$item_oth=='LOCUST'] <- 511
ihs5$item_code[ihs5$item_oth=='ZIWALA'] <- 511
ihs5$item_code[ihs5$item_oth=='MPILU'] <- 403
ihs5$item_code[ihs5$item_oth=='KHUNGUZU'] <- 302
ihs5$item_code[ihs5$item_oth=='MISALE'] <- 108
ihs5$item_code[ihs5$item_oth=='KHOLOWA'] <- 404
ihs5$item_code[ihs5$item_oth=='CHISOSO'] <- 404
ihs5$item_code[ihs5$item_oth=='BONONGWE'] <- 404
ihs5$item_code[ihs5$item_oth=='KHWANYA'] <- 404
ihs5$item_code[ihs5$item_oth=='KWANYA'] <- 404
ihs5$item_code[ihs5$item_oth=='MTAMBE'] <- 404
ihs5$item_code[ihs5$item_oth=='LUNI'] <- 404
ihs5$item_code[ihs5$item_oth=='MPHONDA'] <- 410
ihs5$item_code[ihs5$item_oth=='EGGS PLANTS'] <- 402
ihs5$item_code[ihs5$item_oth=='THERERE'] <- 411
ihs5$item_code[ihs5$item_oth=='CHAMWAMBA'] <- 411
ihs5$item_code[ihs5$item_oth=='CHITAMBE'] <- 404
ihs5$item_code[ihs5$item_oth=='DORGHUM FLOUR'] <- 108
ihs5$item_code[ihs5$item_oth=='GATHERED THERERE'] <- 411
ihs5$item_code[ihs5$item_oth=='KATATA'] <- 411
ihs5$item_code[ihs5$item_oth=='MABILINGANO'] <- 402
ihs5$item_code[ihs5$item_oth=='PUMPKINS'] <- 410
ihs5$item_code[ihs5$item_oth=='THERERE WACHINYOLOMONYA'] <- 411
ihs5$item_code[ihs5$item_oth=='THERERE WAMASAMBA'] <- 411
ihs5$item_code[ihs5$item_oth=='WILD THERERE'] <- 411
ihs5$item_code[ihs5$item_oth=='DUCK'] <- 509
ihs5$item_code[ihs5$item_oth=='ANTELOPE'] <- 505
ihs5$item_code[ihs5$item_oth=='BIRD'] <- 509
ihs5$item_code[ihs5$item_oth=='TSETSENYA (LOOKS LIKE IMSECTS'] <- 511
ihs5$item_code[ihs5$item_oth=='CHIBUKU BEER'] <- 908
ihs5$item_code[ihs5$item_oth=='FIRST CHOICE MILK'] <- 701
ihs5$item_code[ihs5$item_oth=='SODA'] <- 812
ihs5$item_code[ihs5$item_oth=='ROYCO'] <- 811
ihs5$item_code[ihs5$item_oth=='MIXED BEANS'] <- 302
ihs5$item_code[ihs5$item_oth=='BUTTER BEANS'] <- 301
ihs5$item_code[ihs5$item_oth=='RED BEANS'] <- 302
ihs5$item_code[ihs5$item_oth=='YELLOW BEANS'] <- 302
ihs5$item_code[ihs5$item_oth=='ROASTED SWEET POTATO'] <- 832
ihs5$item_code[ihs5$item_oth=='IRISH POTATOES'] <- 205
ihs5$item_code[ihs5$item_oth=='PEPPER'] <- 811
ihs5$item_code[ihs5$item_oth=='YAMS'] <- 208
ihs5$item_code[ihs5$item_oth=='COCOYAMS'] <- 208
ihs5$item_code[ihs5$item_oth=='COOKED SWEET POTATOES (VENDORS)'] <- 831
ihs5$item_code[ihs5$item_oth=='NKHWALI'] <- 509
ihs5$item_code[ihs5$item_oth=='BAKES'] <- 906
ihs5$item_code[ihs5$item_oth=='BAKES JUICE'] <- 906
ihs5$item_code[ihs5$item_oth=='DRAGON  FROZY'] <- 907
ihs5$item_code[ihs5$item_oth=='JOLKY JUS'] <- 906
ihs5$item_code[ihs5$item_oth=='JOLLY JAS JUICE'] <- 906
ihs5$item_code[ihs5$item_oth=='THUMPS UP'] <- 907
ihs5$item_code[ihs5$item_oth=='WAKA JUICE'] <- 906
ihs5$item_code[ihs5$item_oth=='MEXICAN APPLE'] <- 609
ihs5$item_code[ihs5$item_oth=='CUSTARD APPLE'] <- 609
ihs5$item_code[ihs5$item_oth=='LEMON'] <- 603
ihs5$item_code[ihs5$item_oth=='WILD FRUIT'] <- 608
ihs5$item_code[ihs5$item_oth=='MAXICAN APPLE'] <- 609
ihs5$item_code[ihs5$item_oth=='WILD PASSION FRUIT'] <- 608
ihs5$item_code[ihs5$item_oth=='MEAN APPLE'] <- 609
ihs5$item_code[ihs5$item_oth=='CUCUMBER'] <- 409
ihs5$item_code[ihs5$item_oth=='MTAMBE'] <- 404
ihs5$item_code[ihs5$item_oth=='PINEAPPLE'] <- 604

```

>MERGE IN CONVERSION FACTORS

```{r}

#Merge in household geovariables 
ihs5 %>% left_join(., hhfilt, by = c("id" = "case_id"))
ihs5 <- ihs5 %>% left_join(., hhfilt, by = c("id" = "case_id"))


#Merge in conversion factors
convf$item_code <- as.factor(convf$item_code)
convf$region <- as.factor(convf$region)
convf$unit_code <- as.factor(convf$unit_code)
ihs5$region <- as.factor(ihs5$region)

ihs5 %>% 
    left_join(., convf, by=c("region" = "region",
                             "item_code" = "item_code", 
                             "purc_unitA" = "unit_code"))

ihs5 <- ihs5 %>% 
    left_join(., convf, by=c("region" = "region",
                             "item_code" = "item_code", 
                             "purc_unitA" = "unit_code"))

ihs5 <- ihs5 %>% rename(
  purc_item_name_factor = 'item_name',
  purc_factor = 'factor',
  purc_unit_name_factor = 'unit_name',
  purc_Otherunit_factor = 'Otherunit')

ihs5 <- ihs5 %>% select(id:item_name_original, purc_item_name_factor, item_oth, purc_quant:purc_unit_oth, purc_unit_name_factor:purc_factor,  region)

```

>CREATE DATA FRAME OF FOOD ITEMS ONLY CONSUMED BY HOUSEHOLD

```{r}
purc_consed <- ihs5 %>% filter(purc_quant != 0) %>% arrange(id)

#Creating a unique id variable of fcode, ucode and rcode
purc_consed$purc_unique_id <- paste0(as.character(purc_consed$item_code), "_",as.character(purc_consed$purc_unitA), "_", as.character(purc_consed$region))

#Food unit codes in IHS5 without available factors from ihs5 conversion factor database
purc_consed %>% filter(.,is.na(purc_factor))

```


>FILLING IN MISSING FACTORS(NAs) WITH AVAILABLE FACTORS FROM IHS4

```{r}

#--- Read in the file 
ihs4factor <- read_excel ("C:/Users/Gareth Osman/OneDrive - LUANAR/Documents/MAPS Project/ihs5 Cleaning/ihs4factors_v5.xls")
ihs4factor <- ihs4factor %>% select(measure_id:unit, ihs4factor_n:ihs4factor_s)

#Renaming variables 
names(ihs4factor) [names(ihs4factor) == 'ihs4factor_n'] <- '1'
names(ihs4factor) [names(ihs4factor) == 'ihs4factor_c'] <- '2'
names(ihs4factor) [names(ihs4factor) == 'ihs4factor_s'] <- '3'

#Gather wide columns into a longer column by increasing the number of rows and decreasing the number of columns
ihs4factor <- pivot_longer(ihs4factor, cols=6:8, names_to= "Region_ihs4factor", values_to = "ihs4factors", values_drop_na = TRUE)

#Renaming variables
ihs4factor <- ihs4factor %>% rename(
ihs4_fcode = 'fcode',
ihs4_item = 'item',
ihs4_ucode = 'ucode',
ihs4_unit = 'unit')

ihs4factor$ihs4_fcode <- as.factor(ihs4factor$ihs4_fcode)

#creating a unique id of fcode, ucode and rcode in the ihs4factor
 ihs4factor$measure_id <- paste0(as.character(ihs4factor$measure_id), "_", as.character(ihs4factor$Region_ihs4factor))
 
 #creating a data frame of missing conversion factors
 ihs5_factor_missing <- purc_consed %>% filter(.,is.na(purc_factor)) %>% select(id, purc_unique_id, item_code, item_name_original, purc_factor)
 
 ihs5_factor_missing %>% filter(.,is.na(purc_factor)) %>% count(purc_factor)
 
 #merge in ihs5_missing factor file with the ihs4_factors
ihs5_factor_missing <- ihs5_factor_missing %>% left_join(., ihs4factor, by=c('purc_unique_id' = 'measure_id', 'item_code' = 'ihs4_fcode'))

ihs5_factor_missing <- select(ihs5_factor_missing, -5)

#filling in missing factors with available factors from ihs4 using the match function
 purc_consed$purc_factor[is.na(purc_consed$purc_factor)] <- ihs5_factor_missing$ihs4factors[match(purc_consed$purc_unique_id, ihs5_factor_missing$purc_unique_id)][which(is.na(purc_consed$purc_factor))]
 

 purc_consed %>% filter(.,is.na(purc_factor)) %>% count(purc_factor)
 

```

>CREATING A NEW VARIABLE OF QUANTITY IN KG 

the new variable was created by multiplying 'cons_quant' by 'factor'

```{r}

purc_consed <- purc_consed %>% mutate(purc_quant_kg = purc_quant * purc_factor)

write.csv(purc_consed, "C:/Users/Gareth Osman/OneDrive - LUANAR/Documents/MAPS Project/ihs5 Cleaning/ihs5_purchased_clean.csv")

```



