---
title: "summaryMn"
author: "Gareth Osman"
date: "10/6/2021"
output: html_document
---

>Packages 

```{r}
library(epiDisplay)
library(foreign)
library(psych)
library(tidyverse)

```

>Apparent intake/nutrient density inadequacy (9 vitamins)

```{r}
ihs5.va <- read.csv(here::here('va.csv'))
ihs5.vc <- read.csv(here::here('vc.csv'))
ihs5.ve <- read.csv(here::here('ve.csv'))
ihs5.thia <- read.csv(here::here('thia.csv'))
ihs5.ribo <- read.csv(here::here('ribo.csv'))
ihs5.niac <- read.csv(here::here('niac.csv'))
ihs5.vb6 <- read.csv(here::here('vb6.csv'))
ihs5.fol <- read.csv(here::here('fol.csv'))
ihs5.vb12 <- read.csv(here::here('vb12.csv'))



ihs5.va$afe.inad<- as.factor(ihs5.va$afe.inad)
ihs5.vc$afe.inad<- as.factor(ihs5.vc$afe.inad)
ihs5.ve$afe.inad<- as.factor(ihs5.ve$afe.inad)
ihs5.thia$afe.inad<- as.factor(ihs5.thia$afe.inad)
ihs5.ribo$afe.inad<- as.factor(ihs5.ribo$afe.inad)
ihs5.niac$afe.inad<- as.factor(ihs5.niac$afe.inad)
ihs5.vb6$afe.inad<- as.factor(ihs5.vb6$afe.inad)
ihs5.fol$afe.inad<- as.factor(ihs5.fol$afe.inad)
ihs5.vb12$afe.inad<- as.factor(ihs5.vb12$afe.inad)

ihs5.va$den.inad<- as.factor(ihs5.va$den.inad)
ihs5.vc$den.inad<- as.factor(ihs5.vc$den.inad)
ihs5.ve$den.inad<- as.factor(ihs5.ve$den.inad)
ihs5.thia$den.inad<- as.factor(ihs5.thia$den.inad)
ihs5.ribo$den.inad<- as.factor(ihs5.ribo$den.inad)
ihs5.niac$den.inad<- as.factor(ihs5.niac$den.inad)
ihs5.vb6$den.inad<- as.factor(ihs5.vb6$den.inad)
ihs5.fol$den.inad<- as.factor(ihs5.fol$den.inad)
ihs5.vb12$den.inad<- as.factor(ihs5.vb12$den.inad)

ihs5.va %>% 
  group_by(den.inad) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / 11432*100) %>% 
  filter(den.inad=="Inadequate")

 ihs5.va %>% 
    group_by(afe.inad) %>% 
    summarise(n = n()) %>%
    mutate(freq = n / 11432*100) %>% 
    filter(afe.inad=="Inadequate")
  
```

>Apparent intake/nutrient density inadequacy (7 minerals)

```{r}
ihs5.ca <- read.csv(here::here('ca.csv'))
ihs5.cu <- read.csv(here::here('cu.csv'))
ihs5.mg <- read.csv(here::here('mg.csv'))
ihs5.k <- read.csv(here::here('k.csv'))
ihs5.na <- read.csv(here::here('na.csv'))
ihs5.se <- read.csv(here::here('se.csv'))
ihs5.zn <- read.csv(here::here('zn.csv'))

ihs5.ca$afe.inad<- as.factor(ihs5.ca$afe.inad)
ihs5.cu$afe.inad<- as.factor(ihs5.cu$afe.inad)
ihs5.mg$afe.inad<- as.factor(ihs5.mg$afe.inad)
ihs5.k$afe.inad<- as.factor(ihs5.k$afe.inad)
ihs5.na$afe.inad<- as.factor(ihs5.na$afe.inad)
ihs5.se$afe.inad<- as.factor(ihs5.se$afe.inad)
ihs5.zn$afe.inad<- as.factor(ihs5.zn$afe.inad)


ihs5.ca$den.inad<- as.factor(ihs5.ca$den.inad)
ihs5.cu$den.inad<- as.factor(ihs5.cu$den.inad)
ihs5.mg$den.inad<- as.factor(ihs5.mg$den.inad)
ihs5.k$den.inad<- as.factor(ihs5.k$den.inad)
ihs5.na$den.inad<- as.factor(ihs5.na$den.inad)
ihs5.se$den.inad<- as.factor(ihs5.se$den.inad)
ihs5.zn$den.inad<- as.factor(ihs5.zn$den.inad)



ihs5.ca %>% 
  group_by(den.inad) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / 11432*100) %>% 
  filter(den.inad=="Inadequate")

 ihs5.ca %>% 
    group_by(afe.inad) %>% 
    summarise(n = n()) %>%
    mutate(freq = n / 11432*100) %>% 
    filter(afe.inad=="Inadequate") 
  


```


>Probabilistic inadequacy estimates for iron (Mineral)

```{r}
ihs5.fe <- read.csv(here::here('fe.csv'))


#Probabistic AME
fe.inad.prop <- read.csv(here::here('fe.inad.prop.csv'))

colnames(fe.inad.prop)[1] <- "fe.prop.group"


fe.ame.prop <- ihs5.fe %>%
  #filter(urban==1) %>% 
  #filter(res.sep==5) %>%
  mutate(fe.prop.group = case_when(
                                      ai.afe <= 15 ~ 1,
                                      ai.afe <= 16.7 & ai.afe > 15 ~ 2,
                                      ai.afe <= 18.7 & ai.afe > 16.7 ~ 3,
                                      ai.afe <= 21.4 & ai.afe > 18.7 ~ 4,
                                      ai.afe <= 23.6 & ai.afe > 21.4 ~ 5,
                                      ai.afe <= 25.7 & ai.afe > 23.6 ~ 6,
                                      ai.afe <= 27.8 & ai.afe > 25.7 ~ 7,
                                      ai.afe <= 30.2 & ai.afe > 27.8 ~ 8,
                                      ai.afe <= 33.2 & ai.afe > 30.2 ~ 9,
                                      ai.afe <= 37.3 & ai.afe > 33.2 ~ 10,
                                      ai.afe <= 45.0 & ai.afe > 37.3 ~ 11,
                                      ai.afe <= 53.5 & ai.afe > 45.0 ~ 12,
                                      ai.afe <= 63.0 & ai.afe > 53.5 ~ 13,
                                      ai.afe > 63 ~ 14)) %>% 
                                      count(fe.prop.group) %>% mutate(fe.prop = n/11432*100)

fe.ame.prop <- merge(x=fe.ame.prop, y=fe.inad.prop , by.x='fe.prop.group', by.y='fe.prop.group', fill=-9999, all.x = TRUE)
fe.ame.prop <- fe.ame.prop %>% mutate(prev.inad = fe.prop * prob.inad)
sum(fe.ame.prop$prev.inad, na.rm = TRUE)

```

>Probabistic Nutrient Density

```{r}

fe.den.prop <- ihs5.fe %>%
  #filter(urban==1) %>% 
  #filter(res.sep==4) %>%
  mutate(fe.prop.group = case_when(
                                      den <= 15/2150*1000 ~ 1,
                                      den <= 16.7/2150*1000 & den > 15/2150*1000 ~ 2,
                                      den <= 18.7/2150*1000 & den > 16.7/2150*1000 ~ 3,
                                      den <= 21.4/2150*1000 & den > 18.7/2150*1000 ~ 4,
                                      den <= 23.6/2150*1000 & den > 21.4/2150*1000 ~ 5,
                                      den <= 25.7/2150*1000 & den > 23.6/2150*1000 ~ 6,
                                      den <= 27.8/2150*1000 & den > 25.7/2150*1000 ~ 7,
                                      den <= 30.2/2150*1000 & den > 27.8/2150*1000 ~ 8,
                                      den <= 33.2/2150*1000 & den > 30.2/2150*1000 ~ 9,
                                      den <= 37.3/2150*1000 & den > 33.2/2150*1000 ~ 10,
                                      den <= 45.0/2150*1000 & den > 37.3/2150*1000 ~ 11,
                                      den <= 53.5/2150*1000 & den > 45.0/2150*1000 ~ 12,
                                      den <= 63.0/2150*1000 & den > 53.5/2150*1000 ~ 13,
                                      den > 63/2150*1000 ~ 14)) %>% 
                                      count(fe.prop.group) %>% mutate(fe.prop = n/11432*100)

fe.den.prop <- merge(x=fe.den.prop, y=fe.inad.prop , by.x='fe.prop.group', by.y='fe.prop.group', fill=-9999, all.x = TRUE)
fe.den.prop <- fe.den.prop %>% mutate(prev.inad = fe.prop * prob.inad)
sum(fe.den.prop$prev.inad, na.rm = TRUE)

```


#This chunk of codes analyses the top 5 vitamins (ie vitamin C, E, B3, B2, B12) and top 4 minerals (ie. Calcium, Iron, Zinc & Selenium) that have highest percantage of population with inadequancy. 

Apparent consumption quantity for the 5 vitamins
```{r}

ihs5.vc <- read.csv(here::here('vc.csv'))
ihs5.vc <- select(ihs5.vc, -X)

ihs5.ve <- read.csv(here::here('ve.csv'))
ihs5.ve <- select(ihs5.ve, -X)

ihs5.ribo <- read.csv(here::here('ribo.csv'))
ihs5.ribo <- select(ihs5.ribo, -X)

ihs5.niac <- read.csv(here::here('niac.csv'))
ihs5.niac <- select(ihs5.niac, -X)

ihs5.vb12 <- read.csv(here::here('vb12.csv'))
ihs5.vb12 <- select(ihs5.vb12, -X)

ihs5.sep <- read.csv(here::here('ihs5.sep.csv'))
ihs5.sep <- ihs5.sep %>% select(HHID, region, urban, res.sep)


ihs5.vc <- merge(x=ihs5.vc, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.ve <- merge(x=ihs5.ve, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.niac <- merge(x=ihs5.niac, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.ribo <- merge(x=ihs5.ribo, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.vb12 <- merge(x=ihs5.vb12, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)

ihs5.vc %>% 
  filter(ai.afe !=0) %>% 
  group_by(urban) %>% 
  summarise(median = median(ai.afe))

#print( sprintf(a, fmt = '%.4f')  )
```

>Apparent consumption quantity for the 4 minerals

```{r}

ihs5.ca <- read.csv(here::here('ca.csv'))
ihs5.ca <- select(ihs5.ca, -X)

ihs5.fe <- read.csv(here::here('fe.csv'))
ihs5.fe <- select(ihs5.fe, -X)

ihs5.se <- read.csv(here::here('se.csv'))
ihs5.se <- select(ihs5.se, -X)

ihs5.zn <- read.csv(here::here('zn.csv'))
ihs5.zn <- select(ihs5.zn, -X)

ihs5.sep <- read.csv(here::here('ihs5.sep.csv'))
ihs5.sep <- ihs5.sep %>% select(HHID, region, urban, res.sep)


ihs5.ca <- merge(x=ihs5.ca, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.fe <- merge(x=ihs5.fe, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.se <- merge(x=ihs5.se, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.zn <- merge(x=ihs5.zn, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)

ihs5.ca %>% 
  filter(ai.afe !=0) %>% 
  group_by(urban) %>% 
  summarise(median = median(ai.afe))

```

> Apparent intake inadequency for ENERGY 

```{r}
ihs5.ener <- read.csv(here::here('ENER.csv'))
ihs5.ener <- select(ihs5.ener, -X)

ihs5.sep <- read.csv(here::here('ihs5.sep.csv'))
ihs5.sep <- ihs5.sep %>% select(HHID, region, urban, res.sep)

ihs5.ener$afe.inad<- as.factor(ihs5.ener$afe.inad)

ihs5.ener <- merge(x=ihs5.ener, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)

ihs5.ener %>% 
    group_by(afe.inad) %>% 
    summarise(n = n()) %>%
    mutate(freq = n / 11432*100) %>% 
    filter(afe.inad=="Inadequate")

ihs5.ener %>% 
    group_by(afe.inad, region) %>% 
    summarise(n = n()) %>%
    mutate(freq = n / 3951*100) %>% 
    filter(afe.inad=="Inadequate") %>% filter(region==2)

```

>Median apparent consumption quantity for ENERGY 

```{r}

ihs5.ener <- read.csv(here::here('ENER.csv'))
ihs5.ener <- select(ihs5.ener, -X)

ihs5.sep <- read.csv(here::here('ihs5.sep.csv'))
ihs5.sep <- ihs5.sep %>% select(HHID, region, urban, res.sep)

ihs5.ener <- merge(x=ihs5.ener, y=ihs5.sep, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)


ihs5.ener %>% 
  filter(ai.afe !=0) %>% 
  group_by(urban) %>% 
  summarise(median = median(ai.afe))

```

