---
title: "IHS5_cleaning_Step #2"
author: "Gareth Osman"
date: "4/16/2023"
output: html_document
---

```{r}

#Loading package
library(tidyverse)

#--- Change to show no scientific notation & round to 3 decimal places
options(scipen = 10, digits=3) 

#--- Read in the file
ihs5 <- read.csv(here::here('hh_mod_g_initial.csv'))
ihs5 <- select(ihs5, -X)

ihs5factor.other <- read.csv(here::here('ihs5factor.other(23).csv'))
ihs5factor.other <- select(ihs5factor.other, -X)
```
> stage 2: CLEANING UNITS LABELED AS 'OTHER'

#The food conversion factor file contains several food items that share the same unit name but have distinct conversion factors. The other unit is denoted by the code '23', which stands for 'Others'. To avoid duplications during the global database merge, I assigned a unique code in the ihs5 food conversion factor file to each code labeled with '23' in the unit code. Hence this chuck of codes prioritizes the factors for the same unit name that are labelled as 'other'. This is because though the unit name was the same but in reality they were different types of units thats why the 'other unit' which was not part of the standard unit was recorded as 'Other' in the global database. As a result, the cleaning targeted specific food items that had such cases of unit measure.

#Additionally, there were non-standard units such as COB, DOZEN, and TRAY etc included in the conversion factors. Hence, I assigned a unique code to those as well. 


```{r}

ihs5[ihs5$cons_unit_oth == "BASIN", "cons_unitA"] <- '97'
ihs5[ihs5$cons_unit_oth == "CUP", "cons_unitA"] <- '84'
ihs5[ihs5$cons_unit_oth == "PAIL (UNSPECIFIED)", "cons_unitA"] <- '4D'
ihs5[ihs5$cons_unit_oth == "TRAY", "cons_unitA"] <- '75'
ihs5[ihs5$cons_unit_oth == "WHOLE", "cons_unitA"] <- '76'
ihs5[ihs5$cons_unit_oth == "20 LITRE PAIL", "cons_unitA"] <- '4E'
ihs5[ihs5$cons_unit_oth == "COB", "cons_unitA"] <- '77'
ihs5[ihs5$cons_unit_oth == "DOZEN", "cons_unitA"] <- '78'
ihs5[ihs5$cons_unit_oth == "FULL", "cons_unitA"] <- '79'
ihs5[ihs5$cons_unit_oth == "QUARTER", "cons_unitA"] <- '80'
ihs5[ihs5$cons_unit_oth == "10L BUCKET", "cons_unitA"] <- '81'
ihs5[ihs5$cons_unit_oth == "3LITRE BUCKET", "cons_unitA"] <- '82'
ihs5[ihs5$cons_unit_oth == "GABA PAIL", "cons_unitA"] <- '4F'
ihs5[ihs5$cons_unit_oth == "HALF PAIL", "cons_unitA"] <- '4G'

ihs5[ihs5$item_code==102 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==104 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==205 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==306 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==309 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==312 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==414 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==811 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==814 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==817 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==906 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==910 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==5023 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==5033 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'
ihs5[ihs5$item_code==5123 & ihs5$cons_unit_oth=='PIECE', 'cons_unitA'] <- '85'

ihs5[ihs5$item_code==103 & ihs5$cons_unit_oth=='PAIL (SMALL)', 'cons_unitA'] <- '87'
ihs5[ihs5$item_code==203 & ihs5$cons_unit_oth=='PAIL (SMALL)', 'cons_unitA'] <- '87'
ihs5[ihs5$item_code==303 & ihs5$cons_unit_oth=='PAIL (SMALL)', 'cons_unitA'] <- '87'
ihs5[ihs5$item_code==308 & ihs5$cons_unit_oth=='PAIL (SMALL)', 'cons_unitA'] <- '87'
ihs5[ihs5$item_code==408 & ihs5$cons_unit_oth=='PAIL (SMALL)', 'cons_unitA'] <- '87'

ihs5[ihs5$item_code==203 & ihs5$cons_unit_oth=='PAIL (MEDIUM)', 'cons_unitA'] <- '89'
ihs5[ihs5$item_code==408 & ihs5$cons_unit_oth=='PAIL (MEDIUM)', 'cons_unitA'] <- '89'
ihs5[ihs5$item_code==601 & ihs5$cons_unit_oth=='PAIL (MEDIUM)', 'cons_unitA'] <- '89'
ihs5[ihs5$item_code==203 & ihs5$cons_unit_oth=='PAIL', 'cons_unitA'] <- '92'

ihs5[ihs5$item_code==203 & ihs5$cons_unit_oth=='5L BUCKET', 'cons_unitA'] <- '86'
ihs5[ihs5$item_code==204 & ihs5$cons_unit_oth=='5L BUCKET', 'cons_unitA'] <- '86'
ihs5[ihs5$item_code==302 & ihs5$cons_unit_oth=='5L BUCKET', 'cons_unitA'] <- '86'
ihs5[ihs5$item_code==306 & ihs5$cons_unit_oth=='5L BUCKET', 'cons_unitA'] <- '86'
ihs5[ihs5$item_code==311 & ihs5$cons_unit_oth=='5L BUCKET', 'cons_unitA'] <- '86'
ihs5[ihs5$item_code==312 & ihs5$cons_unit_oth=='5L BUCKET', 'cons_unitA'] <- '86'
ihs5[ihs5$item_code==313 & ihs5$cons_unit_oth=='5L BUCKET', 'cons_unitA'] <- '86'
ihs5[ihs5$item_code==408 & ihs5$cons_unit_oth=='5L BUCKET', 'cons_unitA'] <- '86'
ihs5[ihs5$item_code==601 & ihs5$cons_unit_oth=='5L BUCKET', 'cons_unitA'] <- '86'

ihs5[ihs5$item_code==203 & ihs5$cons_unit_oth=='PAIL (LARGE)', 'cons_unitA'] <- '88'
ihs5[ihs5$item_code==606 & ihs5$cons_unit_oth=='PAIL (LARGE)', 'cons_unitA'] <- '88'
ihs5[ihs5$item_code==601 & ihs5$cons_unit_oth=='PAIL (LARGE)', 'cons_unitA'] <- '88'

ihs5[ihs5$item_code==404 & ihs5$cons_unit_oth=='BUNDLE', 'cons_unitA'] <- '91'
ihs5[ihs5$item_code==306 & ihs5$cons_unit_oth=='PLATE', 'cons_unitA'] <- '94'
ihs5[ihs5$item_code==602 & ihs5$cons_unit_oth=='BUNCH SMALL', 'cons_unitA'] <- '93'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='SMALL PACKET', 'cons_unitA'] <- '90'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='SMALL PACKAGE', 'cons_unitA'] <- '95'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='RUBE', 'cons_unitA'] <- '83'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='TBE', 'cons_unitA'] <- '96'
ihs5[ihs5$item_code==306 & ihs5$cons_unit_oth=='90 GRAM','cons_unitA'] <- '99'
ihs5[ihs5$item_code==301 & ihs5$cons_unit_oth=='90 GRAM','cons_unitA'] <- '99'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='100 GRAM','cons_unitA'] <- '99A'
ihs5[ihs5$item_code==829 & ihs5$cons_unit_oth=='MEAL','cons_unitA'] <- '94'
ihs5[ihs5$item_code==829 & ihs5$cons_unit_oth=='MEALS','cons_unitA'] <- '94'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='PACKET MEDIUM','cons_unitA'] <- '99B'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='125G PACKET','cons_unitA'] <- '99C'
ihs5[ihs5$item_code==810 & ihs5$cons_unit_oth=='200 GRAM','cons_unitA'] <- '99D'
ihs5[ihs5$item_code==505 & ihs5$cons_unit_oth=='500 GRAM','cons_unitA'] <- '99E'
ihs5[ihs5$item_code==910 & ihs5$cons_unit_oth=='100 GRAM','cons_unitA'] <- '99A'
ihs5[ihs5$item_code==301 & ihs5$cons_unit_oth=='500 GRAM','cons_unitA'] <- '99E'
ihs5[ihs5$item_code==113 & ihs5$cons_unit_oth=='90 GRAM','cons_unitA'] <- '99'
ihs5[ihs5$item_code==113 & ihs5$cons_unit_oth=='60 GRAMS','cons_unitA'] <- '99F'
ihs5[ihs5$item_code==116 & ihs5$cons_unit_oth=='300GRAMS','cons_unitA'] <- '99G'
ihs5[ihs5$item_code==901 & ihs5$cons_unit_oth=='100 GRAM','cons_unitA'] <- '99A'
ihs5[ihs5$item_code==811 & ihs5$cons_unit_oth=='100 GRAM','cons_unitA'] <- '99A'

```

##CLEANING FUNCTIONS 

Create unique ID variable

```{r}

ihs5$measure_id <- paste0(as.character(ihs5$item_code), "_", as.character(ihs5$cons_unitA))
ihs5$measure_id[is.na(ihs5$cons_unit)] <- NA
 
#Creating data frame of food items consumed by household 
ihs5 <- ihs5 %>% filter(cons_quant !=0) %>% arrange(id)


write.csv(ihs5, here::here('hh_mod_g_clean.csv'))

```






