---
title: "non_standard_units"
author: "Gareth Osman"
date: "5/30/2023"
output: html_document
---

>PACKAGES 

```{r}
#--- Load libraries 
library(readxl)
library(tidyverse)

options(digits=2)
```

>DATA 

```{r}
#--- Read in the files

ihs5 <- read.csv(here::here('hh_mod_g_clean.csv'))
hh_id <- haven::read_dta(here::here('hh_mod_a_filt.dta'))
ihs5factor <- read.csv(here::here('ihs5factors_clean.csv'))
ihs4factor <- read_excel (here::here('ihs4factors_v5.xls'))


```

>CONVERT NON-STANDARD UNITS (NSU) TO KGs

This aims to convert all food consumption quantities reported in NSU into KGs consumed of each food item. 

Background: This R-markdown converts HCES food consumption data from standard (e.g. mL, L, g) and non-standard units (e.g. pail, basin, heaps) to kilograms in order to compare household food consumption. Conversion factors for non-standard units (NSUs) equal the mass (in kg) of one unit of each food item. Hence, a conversion factor is required for each food item-unit combination.

The World Bank's LSMS team provided NSU conversion factors for a variety of different food items, disaggregated by region (bucket/pail sizes vary by region in Malawi). We prioritized these World Bank NSU factors, and while I attempted to retain as much of the World Bank data as possible (i.e. retaining regional factors and applying them according to the region in which the household was located), some cleaning was necessary. For example, our ihs5 factor file contained a number of units labeled '23' but described as existing 'ucodes'. However, distinct 'ucodes' were assigned to '23' (OTHER SPECIFY) in order to distinguish it from known units.

There were numerous food-unit factors that were not included in the World Bank factors for IHS5 (53,776). As a result, they were replaced with readily available IHS4 factors (a file prepared by Kevin Tang [doi: 10.1111/nyas.14697])

> Create  data frame

1. Label households by region
  Food consumption data and geographic region data are located in different parts of the survey. Bring the two pieces of information together and merge.


```{r}
#select variables and change data class of data frames 
ihs5 <- select(ihs5, -X)
ihs5$id <- as.character(ihs5$id)

region <- hh_id %>% select(case_id, region)

ihs4factor <- ihs4factor %>% select(measure_id:unit, ihs4factor_n:ihs4factor_s)
```

> Merge region 

```{r}

ihs5 <- ihs5 %>% left_join(., region, by = c("id" = "case_id"))  %>% arrange(item_code) %>% arrange(HHID)

```

>Merge food-unit specific NSU factor 
  
Merge in the NSU factors according by food item based on what NSU unit each household reported consuming that food item by. Maintain the variation in unit volumes/mass by region.

```{r}

#Change data class 
ihs5factor$item_code <- as.factor(ihs5factor$item_code)
ihs5factor$region <- as.factor(ihs5factor$region)
ihs5factor$unit_code <- as.factor(ihs5factor$unit_code)
ihs5$region <- as.factor(ihs5$region)
ihs5$item_code <- as.factor(ihs5$item_code)


#Creating a unique id of fcode, ucode and rcode
ihs5$measure_id <- paste0(as.character(ihs5$measure_id), "_", as.character(ihs5$region))

ihs5factor$measure_id <- paste0(as.character(ihs5factor$item_code), "_", as.character(ihs5factor$unit_code), "_", as.character(ihs5factor$region))

ihs5factor <- ihs5factor %>% select(measure_id, item_name, unit_name, Otherunit, factor)

ihs5 <- ihs5 %>% left_join(., ihs5factor, by = c("measure_id" = "measure_id")) %>% arrange(id)

ihs5 <- ihs5 %>% rename(item_name_factor = 'item_name')

##Food unit codes in IHS5 without available factors from ihs5 conversion factor database
ihs5 %>% filter(.,is.na(factor)) %>% count()

```

>REPLACE IN MISSING FACTORS(NAs) WITH AVAILABLE IHS4_FACTORS

```{r}

#Renaming variables 
names(ihs4factor) [names(ihs4factor) == 'ihs4factor_n'] <- '1'
names(ihs4factor) [names(ihs4factor) == 'ihs4factor_c'] <- '2'
names(ihs4factor) [names(ihs4factor) == 'ihs4factor_s'] <- '3'


#Gather wide columns into a longer column by increasing the number of rows and decreasing the number of columns
ihs4factor <- pivot_longer(ihs4factor, cols=6:8, names_to= "Region_ihs4factor", values_to = "ihs4factors", values_drop_na = TRUE)


#Renaming variables
ihs4factor <- ihs4factor %>% rename(
 ihs4_fcode = 'fcode',
ihs4_item = 'item',
ihs4_ucode = 'ucode',
ihs4_unit = 'unit')

 
#creating a unique id of fcode, ucode and rcode in the ihs4factor
ihs4factor$measure_id <- paste0(as.character(ihs4factor$measure_id), "_", as.character(ihs4factor$Region_ihs4factor))
 
ihs4factor <- ihs4factor %>% filter(ihs4factors != 0)

#Replace missing factors with available ihs4_factors using the match function
ihs5$factor[is.na(ihs5$factor)] <- ihs4factor$ihs4factors[match(ihs5$measure_id, ihs4factor$measure_id)][which(is.na(ihs5$factor))]

ihs5 %>% filter(.,is.na(factor)) %>% count()

ihs5 %>% group_by(item_name_original, cons_unitA) %>% filter(is.na(factor)) %>% count(factor) %>% arrange(desc(n))

```

>CONVERT NON-STANDARD UNITS TO KGs

Calculate the quantity of each food item consumed in Kg
(Quantity of the food item consumed in the NSU) X (NSU factor)

```{r}

ihs5$cons_kg <- ihs5$cons_quant * ihs5$factor

```

>Convert to quantity consumed per day

IHS5 conducted a 7-day recall. To calculate total consumed per day, just divide by 7. 

```{r}

ihs5$kg_d <- ihs5$cons_kg/7

```

>SAVE 

```{r}

write.csv(ihs5, here::here('hh_mod_g_kg.csv'))

```





