---
title: "food group matching"
author: "Gareth Osman"
date: "9/8/2021"
output: html_document
---

>Packages 

```{r}

library(tidyverse)

```

>Data

```{r}

ihs5 <- read.csv(here::here('hh_mod_g_nep.csv')) #Food consumption data
ihs5 <- select(ihs5, -X)

foodgroup <- read.csv(here::here('FAO.12fg.csv')) #Food groups 
foodgroup <- select(foodgroup, -X)

```

>Labeling food groups of food items

We need to add a variable to group items by the food group that a food item fell under according to the FAO food groups for Household Dietary Diversity. There are 12 food groups in total: 
-Cereals
-Vegetables 
-Milk & milk products
-Fats & oils 
-Legumes & nuts
-Fruits
-Eggs
-Meat
-Fish 
-Sugary foods/beverages 
-Roots & tubers
-Miscellaneous

```{r}

ihs5 <- merge(x=ihs5, y=foodgroup, by.x='item_code', by.y='code', fill=-9999, all.x = TRUE) %>% arrange(item_code) %>% arrange(HHID)

ihs5 <- ihs5 %>% relocate("item_code", .after = g100_d_nep)
ihs5 <- ihs5 %>% rename(food.group ='Category')

```

>DONE : ARCHIVE 

```{r}

write.csv(ihs5, here::here('hh_mod_g_nepFinal.csv'))

```
